---
title: "PIAAC"
author: "EC1027 --- Econometría I"
date: "Curso 2021-2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

# Introducción 

El Programa para la Evaluación Internacional de las Competencias de los adultos (PIAAC) es una iniciativa de la OCDE para valorar el nivel de conocimientos y el grado y distribución de destrezas de la población adulta a lo largo de su vida laboral, entre los 16 y los 65 años. Esta evaluación se encuadra en un marco comparativo coherente para poder establecer contrastes internacionales.

La [página](https://www.oecd.org/skills/piaac/) de la OCDE del PIAAC contiene mucha información. En concreto, proporciona acceso a:

- Las [bases de datos públicas](https://www.oecd.org/skills/piaac/data/) en varios formatos.
- Informes técnicos y los diferentes [informes nacionales](https://www.oecd.org/skills/piaac/country-specific-material.htm).


# Variables 

En el archivo comprimido `esp.csv.gz` se ha recopilado parte de la información para España. En concreto, contiene las siguientes variables:

- `SEQID`: Sequential ID (randomly derived).
- `GENDER_R`: Person resolved gender from BQ and QC check (derived).

  Value   Label
  ------- --------
  1       Male
  2       Female

- `AGE_R`:   Person resolved age from BQ and QC check (derived).
- `AGEG5LFS`:   Age groups in 5-year intervals based on LFS groupings (derived).

  Value   Label
  ------- ------------
  1       Aged 16-19
  2       Aged 20-24
  3       Aged 25-29
  4       Aged 30-34
  5       Aged 35-39
  6       Aged 40-44
  7       Aged 45-49
  8       Aged 50-54
  9       Aged 55-59
  10      Aged 60-65

- `EDCAT7`:   Highest level of formal education obtained (7 categories - derived).

  Value   Label
  ------- ---------------------------------------------------------
  1       Primary or less (ISCED 1 or less)
  2       Lower secondary (ISCED 2, ISCED 3C short)
  3       Upper secondary (ISCED 3A-B, C long)
  4       Post-secondary, non-tertiary (ISCED 4A-B-C)
  5       Tertiary - professional degree (ISCED 5B)
  6       Tertiary - bachelor degree (ISCED 5A)
  7       Tertiary - master/research degree (ISCED 5A/6)
  8       Tertiary - bachelor/master/research degree (ISCED 5A/6)

- `EARNHR`:   Hourly earnings excluding bonuses for wage and salary earners (derived).
- `YRSQUAL`:   Highest level of education obtained imputed into years of education (derived).
- `YRSGET`:   Imputed years of formal education needed to get the job (self-reported - derived).
- `C_Q07_T`:   Current work situation (Trend-IALS/ALL).

  Value   Label
  ------- -----------------------------------
  1       Employed or self employed
  2       Retired
  3       Not working and looking for work
  4       Student (including work programs)
  5       Doing unpaid household work
  6       Other

- `C_D05`:   Current status/work history - Employment status (DERIVED BY CAPI).

  Value   Label
  ------- -------------------------
  1       Employed
  2       Unemployed
  3       Out of the labour force
  4       Not known

- `C_Q09`:   Current status/work history - Years of paid work during lifetime.
- `D_Q03`:   Current work - Economic sector.

  Value   Label
  ------- -------------------------------------------------------------------------------------------------------
  1       The private sector (for example a company)
  2       The public sector (for example the local government or a state school)
  3       A non-profit organisation (for example a charity, professional association or religious organisation)

- `D_Q04`:   Current work - Employee or self-employed.

  Value   Label
  ------- ---------------
  1       Employee
  2       Self-employed

- `D_Q05a1_C`:   Current work - Start of work for employer - Age (categorised, 9 categories).

  Value   Label
  ------- --------------------
  1       Aged 19 or younger
  2       Aged 20-24
  3       Aged 25-29
  4       Aged 30-34
  5       Aged 35-39
  6       Aged 40-44
  7       Aged 45-49
  8       Aged 50-54
  9       Aged 55 or older

- `D_Q05a2`:   Current work - Start of work for employer - Year.
- `D_Q05b2`:   Current work - Start of work for business - Year.
- `D_Q09`:   Current work - Type of contract.

  Value   Label
  ------- --------------------------------------------
  1       An indefinite contract
  2       A fixed term contract
  3       A temporary employment agency contract
  4       An apprenticeship or other training scheme
  5       No contract
  6       Other

- `D_Q12a`:   Current work - Requirements - Education level

  Value   Label
  ------- -------------------------------------------------------------------------
  1       No formal qualification or below ISCED 1
  2       ISCED 1
  3       ISCED 2
  4       ISCED 3C shorter than 2 years
  5       ISCED 3C 2 years or more
  6       ISCED 3A-B
  7       ISCED 3 (without distinction A-B-C, 2y+)
  8       ISCED 4C
  9       ISCED 4A-B
  10      ISCED 4 (without distinction A-B-C)
  11      ISCED 5B
  12      ISCED 5A, bachelor degree
  13      ISCED 5A, master degree
  14      ISCED 6
  15      ISCED 5A bachelor degree, 5A master degree, and 6 (without distinction)

- `D_Q12b`:   Current work - Requirements - To do the job satisfactorily

  Value   Label
  ------- -----------------------------------
  1       This level is necessary
  2       A lower level would be sufficient
  3       A higher level would be needed

- `D_Q12c`:   Current work - Requirements - Related work experience

  Value   Label
  ------- -------------------
  1       None
  2       Less than 1 month
  3       1 to 6 months
  4       7 to 11 months
  5       1 or 2 years
  6       3 years or more

- `J_Q02a`:   Background - Living with spouse or partner.

  Value   Label
  ------- -------
  1       Yes
  2       No

- `J_Q02c`:   Background - Work situation of spouse or partner.

  Value   Label
  ------- ------------------------------------------------------------
  1       Full-time employed (self-employed, employee)
  2       Part-time employed (self-employed, employee)
  3       Unemployed
  4       Pupil, student
  5       Apprentice, internship
  6       In retirement or early retirement
  7       Permanently disabled
  8       In compulsory military or community service
  9       Fulfilling domestic tasks or looking after children/family
  10      Other

- `J_Q03a`:   Background - Children.

  Value   Label
  ------- -------
  1       Yes
  2       No

- `J_Q03b`:   Background - Number of children.
- `J_Q04a`:   Background - Born in country.

  Value   Label
  ------- -------
  1       Yes
  2       No

- `J_Q04c2`:   Background - Year of immigration.
- `J_Q06a_T`:   Mother or female guardian born in country (Trend-IALS/ALL).

  Value   Label
  ------- ----------------
  1       Yes
  2       No
  3       Not applicable

- `J_Q06b_T`:   Highest level of education - mother or female guardian (Trend-IALS/ALL).

  Value   Label
  ------- ------------------------------------
  1       ISCED 1, 2, and 3C short
  2       ISCED 3 (excluding 3C short) and 4
  3       ISCED 5 and 6
  4       Not definable

- `J_Q07a_T`:   Father or male guardian born in country (Trend-IALS/ALL).

  Value   Label
  ------- ----------------
  1       Yes
  2       No
  3       Not applicable

- `J_Q07b_T`:   Highest level of education - father or male guardian (Trend-IALS/ALL).

  Value   Label
  ------- ------------------------------------
  1       ISCED 1, 2, and 3C short
  2       ISCED 3 (excluding 3C short) and 4
  3       ISCED 5 and 6
  4       Not definable

- `J_Q08`:   Background - Number of books at home.

  Value   Label
  ------- ---------------------
  1       10 books or less
  2       11 to 25 books
  3       26 to 100 books
  4       101 to 200 books
  5       201 to 500 books
  6       More than 500 books

- `PVLIT1`:   Literacy scale score - Plausible value 1.
- `PVNUM1`:   Numeracy scale score - Plausible value 1.
- `SPFWT0`:   Final full sample weight.

# Ejemplo

Si el archivo `esp.csv.gz` está en el directorio de trabajo, la base de datos puede leerse con la instrucción:
```{r}
esp <- read.csv("esp.csv.gz")
```

Creamos una nueva base de datos que contiene sólo el logaritmo de los salarios por hora:
```{r}
db <- data.frame(lwage = log(esp$EARNHR))
```

Añadimos la variable ficticia `female` que toma el valor 1 para las mujeres.
```{r}
db$female <- as.integer(esp$GENDER_R == 2)
```

Eliminamos las observaciones incompletas (en este caso se trata en su mayoría de personas que no perciben salarios: desempleados, inactivos y autónomos):

```{r}
db_emp <- na.omit(db)
```

Estimación de la brecha salarial mediante una regresión simple:
```{r}
mod1 <- lm(lwage ~ female, data = db_emp)
summary(mod1)
```

